\section{Tubular neighbourhoods of neural networks}\label{sec:tubular-neural}
Artificial neural networks, or multilayer perceptrons, with common continuous activation functions are examples of Pfaffian functions~\cite{karpinski1997polynomial, bianchini2014complexity}. When used as classifiers, neural networks
subdivide the input space into decision regions separated by semi-Pfaffian decision boundaries and the Pfaffian volume bound in Theorem~\ref{thm:prob_bound} can be used to quantify the probability that a randomly chosen input is close to that boundary. 
The general volume bound for Pfaffian functions is, however, overly pessimistic for neural networks. In this section we derive an improved volume bound for one-layer neural networks with sigmoid activation.

\subsection{Neural networks as Pfaffian functions}
 Let $\cN$ be a fully connected neural network with $\ell$ hidden layers and $h$ hidden units. Such a network is characterized by a function $F=F^{\ell+1}\circ \cdots \circ F^1\colon \R^n\to \R^m$, where
\begin{align*}
  F^{i}(x) &= \sigma^{i}(A^{i}x+b^{i}), \quad i\in [\ell],\\
  F^{\ell+1}(x) &= g(x)
\end{align*}
%\begin{align*}
%  &  z^{(0)} = x \\
%  &  z^{(i)} = \sigma (A_i z^{(i-1)} + b^{(i)}), \ 1 \le i \le \ell \\
%  & f(x) = A_{\ell +1} z^{(l)},
%\end{align*}  
for matrices $A^{i} \in \R^{n_{i} \times n_{i-1}}$ (for this to work, we need $n_0 = n$ and $n_{\ell + 1} = m$), vectors $b^{i} \in \R^{n_i}$, and functions $\sigma^i\colon \R^{n_i}\to \R^{n_i}$, where $\sigma^i=(\sigma^i_1,\dots,\sigma^i_{n_i})^{\trans}$ consists of activation functions. 
The function $g\colon \R^{n_\ell}\to \R^m$ is an output function; examples are a linear map $g(x)=A^{\ell+1}x$ or the softmax function (see Example~\ref{ex:softmax}). The number of hidden units is given as the sum $h:=\sum_{i=1}^{\ell} n_i$. 

\begin{figure}[h!]
\centering
\begin{tikzpicture}[thick, scale=1.5]
  % Node placement
  \node [every neuron] (I1) at (-2,0.5) {};
  \node [every neuron] (I2) at (-2,-0.5) {};
  \node [every neuron] (M1) at (-1,1) {};
  \node [every neuron] (M2) at (-1,0) {};
  \node [every neuron] (M3) at (-1,-1) {};
  \node [every neuron] (O1) at (0,0.5) {};
  \node [every neuron] (O2) at (0,-0.5) {};
  
  \node (OO1) at (0.75,0.5) [] {$F_1(x)$};
  \node (OO2) at (0.75,-0.5) [] {$F_2(x)$};
  \node (II1) at (-3,0.5) [] {$x_1$};
  \node (II2) at (-3,-0.5) [] {$x_2$};
  \node [every neuron] (II1) at (-3,0.5) {};
  \node [every neuron] (II2) at (-3,-0.5) {};
  
  % Arrows
  %\draw[->] (III1) -- (II1);
  %\draw[->] (III2) -- (II2);
  \draw[->] (II1) -- (I1);
  \draw[->] (II1) -- (I2);
  \draw[->] (II2) -- (I1);
  \draw[->] (II2) -- (I2);
  \draw[->] (I1) -- (M1);
  \draw[->] (I1) -- (M2);
  \draw[->] (I2) -- (M2);
  \draw[->] (I2) -- (M3);
  \draw[->] (I1) -- (M3);
  \draw[->] (I2) -- (M1);
  \draw[->] (M1) -- (O1);
  \draw[->] (M2) -- (O1);
  \draw[->] (M2) -- (O2);
  \draw[->] (M3) -- (O2);
  \draw[->] (M1) -- (O2);
  \draw[->] (M3) -- (O1);
  \draw[->] (O1) -- (OO1);
  \draw[->] (O2) -- (OO2);
\end{tikzpicture}
\caption{A fully connected neural network with $n = 2, m=2, \ell = 2, h = 5$}\label{fig:nn}
\end{figure}

%\begin{figure}
%    \centering
%          \begin{neuralnetwork}[height=6]
%            \newcommand{\x}[2]{$x_#2$}
%            \newcommand{\y}[2]{$f(x)$}
%            \newcommand{\hfirst}[2]{\small $z^{(1)}_#2$}
%            \newcommand{\hsecond}[2]{\small $z^{(2)}_#2$}
%            \inputlayer[count=3, bias=false, title=Input\\layer, text=\x]
%            \hiddenlayer[count=5, bias=false, title=Hidden\\layer 1, text=\hfirst] \linklayers
%            \hiddenlayer[count=5, bias=false, title=Hidden\\layer 2, text=\hsecond] \linklayers
%            \outputlayer[count=1, title=Output\\layer, text=\y] \linklayers
%        \end{neuralnetwork}
%    \caption{Neural network with $n = 3, \ell = 2, h = 10$}
%\end{figure}

Suppose that the activation functions $\sigma^i_j$ are autonomous Pfaffian functions with format $(\alpha, \beta, s)$ with $s\geq 1$.
That is, each $\sigma^i_j$ can be written as
\begin{equation*}
\sigma^i_j(x) = Q^{ij}(q_1^{ij}(x), \dotsc, q^{ij}_{s}(x)),
\end{equation*}
where $Q^{ij}\in \R[Y_1,\dots,Y_{s}]$ with $\deg Q^{ij}\leq \beta$ and $\boldq^{ij} = (q_1^{ij}, \dotsc, q^{ij}_{s})$ is the Pfaffian chain corresponding to $\sigma^i_j$, satisfying
\begin{equation*}
\frac{\diff{q_r^{ij}}}{\diff{x}} = P_r^{ij} (q_1^{ij}(x), \dotsc, q_r^{ij}(x)),
\end{equation*}
for polynomials $P_r^{ij}$ satisfying $\deg P_r^{ij} \le \alpha$.

If the output function $g$ is also Pfaffian, then Lemma~\ref{le:composition} implies that each component of $F^i$ is a Pfaffian function, with a Pfaffian chain consisting of $s$ functions for every unit in the $i$-th layer of the network. If we denote the output of the $j$-th node in the $i$-th layer by $z^i_j$, i.e., the $j$-th component of the vector $z^i := F^i\circ \cdots \circ F^1(x)$, then this node
contributes the functions
\begin{equation*}
  (q_1^{ij}(z^i_j),\dots,q_s^{ij}(z^i_j))
\end{equation*}
to the Pfaffian chain of the network. We can use Lemma~\ref{le:composition} (2) to derive the following result on the Pfaffian structure of such a neural network. The proof is a simple induction on the depth of the network.

\begin{proposition}\label{prop:nn-format}
Let $F\colon\R^n\to \R^m$ be a function implemented by a neural network with $\ell$ hidden layers and $n_i$ units in hidden layer $i$, for $i\in [\ell]$. Assume that the activation functions at each layer are autonomous Pfaffian functions with format $(\alpha,\beta,s)$, $s\geq 1$. Then for $i\in [\ell]$, the function $F^i\circ \cdots \circ F^1\colon \R^n\to \R^{n_i}$ is Pfaffian with format
\begin{equation*}
  \left(i(\alpha+\beta-1) -\beta+1 , \beta, s\sum_{j=1}^i n_j\right).
\end{equation*}
In particular, at the last hidden layer ($i=\ell$), the format of $F^\ell\circ \cdots \circ F^1$ is
\begin{equation*}
  \left(\ell(\alpha+\beta-1)-\beta+1,\; \beta,\; sh\right),
\end{equation*}
where $h=\sum_{j=1}^\ell n_j$ is the total number of hidden units. 
\end{proposition}

Thus, for a fixed set of activation functions (and hence fixed $\alpha$, $\beta$, $s$), the chain-degree grows linearly in the depth $\ell$, the function degree is constant, and the order grows linearly in $h$.

\begin{remark}\label{re:chain-length-tight}
The order $sh$ is essentially optimal for generic fully connected networks: the $h$ intermediate activation values are algebraically independent functions of $x$ (for generic weights), and each must appear as a chain element since it arises independently in the partial derivatives of the network output via the chain rule. The structural optimization provided by Lemma~\ref{le:composition}(3), which keeps the order additive across layers rather than multiplicative, is already incorporated into the bound.
\end{remark}

\begin{example}[Activation functions]\label{ex:activations}
Table~\ref{tab:activations} lists common activation functions and their Pfaffian formats. Here $\sigma(x)=(1+\econst^{-x})^{-1}$ is the logistic sigmoid, $\phi$ and $\Phi$ denote the standard Gaussian density and cdf, respectively, and $\mathrm{sp}(x)=\ln(1+\econst^x)$ is the softplus function~\cite{glorot2011deep}. GELUs were introduced in~\cite{hendrycks2016gaussian}. Nonsmooth activation functions such as ReLU do not fall into our framework, nor does the ELU, which is not analytic at the origin.

\begin{table}[h!]
  \centering
  \begin{tabular}{lllcccc}
    \toprule
    Activation & $\sigma(x)$ & Chain $\boldq$ & $\alpha$ & $\beta$ & $s$ & Auton.\\
    \midrule
    Exponential & $\econst^x$ & $(\econst^x)$ & 1 & 1 & 1 & Yes \\
    Sigmoid & $(1+\econst^{-x})^{-1}$ & $(\sigma)$ & 2 & 1 & 1 & Yes \\
    Tanh & $\tanh(x)$ & $(\tanh)$ & 2 & 1 & 1 & Yes \\
    Softplus & $\ln(1+\econst^x)$ & $(\sigma,\mathrm{sp})$ & 2 & 1 & 2 & Yes \\
    SiLU/Swish & $x\sigma(x)$ & $(\sigma)$ & 2 & 2 & 1 & No \\
    GELU & $x\Phi(x)$ & $(\phi,\Phi)$ & 2 & 2 & 2 & No \\
    Mish & $x\tanh(\mathrm{sp}(x))$ & $(\sigma,\mathrm{sp},\tanh(\mathrm{sp}))$ & 3 & 2 & 3 & No \\
    Gaussian & $\econst^{-x^2}$ & $(\econst^{-x^2})$ & 2 & 1 & 1 & No \\
    Arctan & $\arctan(x)$ & $((1+x^2)^{-1},\arctan)$ & 3 & 1 & 2 & No \\
    \bottomrule
  \end{tabular}
  \caption{Pfaffian formats of common activation functions.}
  \label{tab:activations}
\end{table}
\end{example}

\begin{example}[Softmax]\label{ex:softmax}
A common output function for classification problems is the softmax function. Each component of the softmax function
\begin{equation}\label{eq:softmax}
  g(x) = \left(\frac{\mathrm{e}^{x_1}}{\sum_{j}\mathrm{e}^{x_j}}, \dots, \frac{\mathrm{e}^{x_m}}{\sum_{j}\mathrm{e}^{x_j}}\right).
\end{equation}
is Pfaffian with format $(2,1,m)$, but the Pfaffian chains of the different components differ by a permutation. For example, 
\begin{equation*}
  \frac{\partial g_i}{\partial x_j}(x) = \begin{cases}
    -g_i(x)g_j(x) & \text{ if } i\neq j\\
    g_i(x)(1-g_i(x)) & \text{ if } i=j.
  \end{cases}
\end{equation*}
In this case, a Pfaffian chain is arrived at by listing the components so that $g_i$ comes last.
\end{example}

When using a neural network as a classifier, we can interpret $F_i(x)$ as representing the likelihood that $x$ belongs to class $i$ (for example, if the output map $g$ is the softmax function). Thus a data point $x\in \R^n$ is assigned to class $j$ if $F_j(x)>F_i(x)$ for all $i\neq j$, with an arbitrary tie break if $F_i(x)=F_j(x)$.
For $i\neq j$, define the functions
$g_{ij} := F_j-F_i$. The map $F$ induces a subdivision of $\R^n$ into $m$ regions
\begin{equation*}
 C_j = \{x\in \R^n \colon g_{ij}(x) \geq 0 \text{ for } i\neq j\}, \quad \quad j\in [m].
\end{equation*}
If $C_j$ has non-empty interior, then the interior $\inter(C_j)$ is the set of $x$ that are unequivocally assigned to class $j$. 
The boundary of $C_j$ and the decision boundary of the classifier are given by
\begin{equation*}
 \Sigma_j = \bigcup_{i\neq j} (C_j\cap C_i), \quad j\in [m] \quad \text{ and }  \quad \Sigma = \bigcup_j \Sigma_j,
\end{equation*}
respectively. 

The following result is a straightforward consequence of the algebraic properties of Pfaffian functions and the definition of semi-Pfaffian sets.

\begin{proposition}\label{prop:decision-pfaffian}
Assume the neural network classifier implemented by $F\colon \R^n\to \R^m$ is a Pfaffian function. Then the functions $g_{ij}:=F_j-F_i$ are also Pfaffian, and the decision boundaries $\Sigma_j$ and $\Sigma$ are semi-Pfaffian sets. 
\end{proposition}

\begin{example}
Consider an $\ell$-layer neural network for classification in which all the activation functions at hidden nodes are of format $(2,1,1)$ (for example, $\tanh$ and the logistic sigmoid $\sigma$). Then the $g_{ij}(x)$ are Pfaffian with format
\begin{equation*}
  (2\ell+1, 1, h+2),
\end{equation*}
where $h$ is the total number of hidden nodes.
\end{example}

\subsection{The Gauss map of a neural network}
The tubular neighbourhood bound for general Pfaffian functions involves a constant that is exponential in the length $s$ of the Pfaffian chain. In the case of neural networks, the chain consists of one element for every hidden unit, so $s=h$ and this factor becomes prohibitive. We show that, for decision boundaries of sigmoid classifiers, the bound can be improved to a polynomial in the number of hidden units.

The strategy is to replace the Khovanski\u i--Rolle bound with a global argument based on the Bernstein--Kushnirenko--Khovanski\u i (BKK) theorem~\cite{bernstein1975} (see also~\cite[Chapter 7,\S 5]{cox2005using}). After an exponential substitution, the Gauss-map system becomes a Laurent polynomial system whose $n$ equations share a single Newton polytope, a zonotope $\conv(A)$ determined by the (integer-scaled) weight vectors. Bernstein's theorem then bounds the number of solutions by $n!\,\operatorname{Vol}(\conv(A))$, which is polynomial in the width. For networks with $\ell\geq 2$ hidden layers, the direct BKK approach is obstructed by the transcendentality of sigmoid composition, and we state the corresponding polynomial-in-widths bound as a conjecture.


\begin{proposition}\label{prop:single-layer-degree}
  Let $f\colon \R^n \to \R$ be given by
  \begin{equation*}
  f(x) = c_0+\sum_{k=1}^n d_k \sigma(a_k^{\trans}x+b_k), 
  \end{equation*}
  where $\sigma$ is the logistic sigmoid, $c_0,b_k\in \R$ and the $a_k\in \mathbb{Q}^{n}$ are in general position with common denominator $q\in\N_{>0}$. Set $V=\{x\colon f(x)=0\}$ and $L := q\cdot\max_{k,i}|a_{ki}|$. 
  Then $V$ is a smooth hypersurface with maximum degree
  \begin{equation}\label{eq:single-layer-bound}
    \md(V) \;\leq\; C(n,L)\cdot w^{n},
  \end{equation}
  where $C(n,L)\leq 2\cdot n!\,(2L)^n$ depends only on $n$ and $L$.
\end{proposition}

\begin{proof}
By Remark~\ref{re:affine}, we may assume the regular value of the Gauss map is $v=e_n$, so that $\md(V)$ is bounded (up to the factor $2$ for the two normal orientations) by the number of solutions of
\begin{equation}\label{eq:single-layer-system}
  f(x) = 0, \qquad \frac{\partial f}{\partial x_j}(x) = 0, \quad j = 1,\ldots, n-1,
\end{equation}
where
\begin{equation}\label{eq:single-layer-f}
  f(x) = c_0 + \sum_{k=1}^w d_k\,\sigma(a_k^{\trans}x + b_k),
  \qquad
  \frac{\partial f}{\partial x_j}(x) = \sum_{k=1}^w d_k\,a_{kj}\,\sigma'(a_k^{\trans}x + b_k).
\end{equation}
Set $\vec p_k := q\,a_k\in\Z^n$, so that $\|\vec p_k\|_\infty\leq L$, and set $y_j=\econst^{-x_j/q}$. We get
\begin{equation*}
  \econst^{-(a_k^{\trans}x + b_k)}
  \;=\; \econst^{-b_k}\,\prod_{j=1}^n \econst^{-(qa_{kj})\,x_j/q}
  \;=\; \econst^{-b_k}\, y^{\vec p_k}
  \;=:\; u_k(y),
\end{equation*}
where we use the notation $y^{\vec p_k}=y_1^{qa_{k1}}\cdots y_n^{qa_{kn}}$. 
Each sigmoid becomes $\sigma_k = (1+u_k)^{-1}$ and its derivative $\sigma'_k = u_k(1+u_k)^{-2}$. Set $\Phi(y):=\prod_{k=1}^w(1+u_k(y))^2$, a strictly positive analytic function on $\R_{>0}^n$. Multiplying~\eqref{eq:single-layer-system} by the non-vanishing factor $\Phi(y)$, the system transforms into the following polynomial system on $(\R_{>0})^n$:
\begin{equation}\label{eq:laurent-system}
  F_0(y) \;:=\; c_0\prod_{k=1}^w(1+u_k)^2 + \sum_{k=1}^w d_k\, \prod_{l\neq k}(1+u_l)^2 \;=\; 0,
\end{equation}
\begin{equation}\label{eq:laurent-system-j}
  F_j(y) \;:=\; \sum_{k=1}^w d_k\,a_{kj}\,u_k\!\!\prod_{l\neq k}(1+u_l)^2 \;=\; 0, \qquad j=1,\ldots,n-1.
\end{equation}
Each factor $(1+u_k)^2 = 1 + 2u_k + u_k^2$ contributes monomials withexponents in $\{0,\vec p_k,2\vec p_k\}$, so all monomials appearing in~\eqref{eq:laurent-system}--\eqref{eq:laurent-system-j} have exponent vectors in the Minkowski sum
\begin{equation}\label{eq:zonotope-support}
  A \;:=\; \sum_{k=1}^w \{0,\vec p_k,2\vec p_k\} \;\subset\; \Z^n.
\end{equation}
Note that \emph{both} $F_0$ and the $F_j$ have support inside the same set $A$: the constant term $c_0$ in $f$ multiplies the full product $\Phi(y)$, while in each $F_j$ the missing constant term simply drops that summand, and all remaining summands have exponents in $A$. The Newton polytopes $P_0, P_1, \ldots, P_{n-1}$ of the $F_i$ are therefore all contained in $\conv(A)$.

By Bernstein's theorem~\cite{bernstein1975}, a system of $n$ Laurent polynomials in $n$ variables with Newton polytopes $P_0,\ldots,P_{n-1}$ has at most
\begin{equation*}
  n!\,\operatorname{MV}(P_0,\ldots,P_{n-1})
\end{equation*}
isolated solutions in $(\C^*)^n$, counted with multiplicity, where $\operatorname{MV}$ denotes the mixed volume, provided no solution escapes to a lower-dimensional face (a non-degeneracy condition discussed later). Since $P_j\subseteq\conv(A)$ for all $j$ and the mixed volume is monotone in each argument, this is at most $n!\,\operatorname{Vol}(\conv(A))$. The diffeomorphism $\Psi\colon x_j\mapsto y_j=\econst^{-x_j/q}$ embeds $\R^n$ bijectively onto $(\R_{>0})^n\subset(\C^*)^n$, and multiplication by $\Phi(y)$ is non-vanishing there, so every isolated real solution of~\eqref{eq:single-layer-system} corresponds to an isolated complex solution of~\eqref{eq:laurent-system}--\eqref{eq:laurent-system-j} in $(\C^*)^n$. Hence the real count is bounded by $n!\,\operatorname{Vol}(\conv(A))$.

The set $A$ lies in the zonotope $Z := \sum_{k=1}^w[0,2\vec p_k]$, so $\conv(A)\subseteq Z$. The zonotope volume formula (see~\cite[Ch.~7]{ziegler1995}) gives
\begin{equation}\label{eq:zonotope-volume}
  \operatorname{Vol}(Z) \;=\; 2^n \sum_{\substack{S\subset[w]\\ |S|=n}} \bigl|\det(\vec p_{k_1},\ldots,\vec p_{k_n})\bigr|,
\end{equation}
where $S=\{k_1,\ldots,k_n\}$. Hadamard's inequality together with $\|\vec p_k\|_\infty\leq L$ gives 
\begin{equation*}
|\det(\vec p_{k_1},\ldots,\vec p_{k_n})|\leq n^{n/2}\,L^n \leq n!\,L^n,
\end{equation*}
and there are $\binom{w}{n}\leq w^n/n!$ subsets, so
\begin{equation*}
  \operatorname{Vol}(\conv(A)) \;\leq\; \operatorname{Vol}(Z) \;\leq\; 2^n\cdot n!\,L^n\cdot\frac{w^n}{n!} \;=\; (2L)^n\,w^n.
\end{equation*}
The Bernstein bound implies that there are at most $n!\,(2L)^n\,w^n$ solutions of~\eqref{eq:single-layer-system}, and the orientation factor $2$ is absorbed into the constant $C(n,L)\leq 2\cdot n!\,(2L)^n$ in~\eqref{eq:single-layer-bound}.

Bernstein's bound is achieved generically and can be exceeded only at degenerate systems. The coefficients of~\eqref{eq:laurent-system}--\eqref{eq:laurent-system-j} are polynomials in the data $(d_k,b_k,c_0,a_{kj})$, so the degenerate locus is a proper algebraic subset of parameter space. Under the general-position hypothesis on $A$ and a Sard-type genericity assumption on $(b_k,d_k,c_0)$, which holds outside a measure-zero set and which may be imposed by an arbitrarily small perturbation that does not increase the zero count, the system avoids the degenerate locus and Bernstein's bound applies.
\end{proof}

\begin{remark}\label{rem:single-layer-lowerbound}
  The exponent $n$ in the bound $\md(V) = O(w^n)$ of Proposition~\ref{prop:single-layer-degree} is tight: a matching lower bound $\md(V) = \Omega(w^n)$ is achieved by a grid construction, as follows. Assume $w = nm$ for some integer $m\ge 2$. For $j\in [n]$ and $k\in[m]$, take weight vectors $a_{j,k} \in \mathbb{Q}^n$ obtained by small rational perturbation of $e_j$, chosen so that the $w$ rows $\{a_{j,k}\}$ are in general position, biases $b_{j,k} = -k$, output weights $d_{j,k} = (-1)^k$, and offset $c_0 = 1/2$. Consider the sharpened family
  \begin{equation*}
    f_\lambda(x) \;=\; \frac{1}{2} + \sum_{j=1}^n\sum_{k=1}^m (-1)^k\,\sigma\bigl(\lambda(a_{j,k}^{\trans}x-k)\bigr), \qquad \lambda \to \infty.
  \end{equation*}
  In the threshold limit $\lambda\to\infty$, each sigmoid converges pointwise to an indicator, and $f_\lambda$ converges to a piecewise constant function $g$ on the cells of the (nearly axis-aligned) hyperplane arrangement. On the cell indexed by $(i_1,\ldots,i_n)\in\{0,1,\ldots,m\}^n$ (with $i_j$ counting how many of the $j$-direction hyperplanes lie ``below'' the cell), $g$ takes the value $c_0 + \sum_{j=1}^n S_{i_j}$, where $S_i = \sum_{k\le i}(-1)^k \in\{0,-1\}$.
  With $c_0 = 1/2$, all-even cells (every $i_j$ even) have $g = +1/2$, and one-odd cells (exactly one $i_j$ odd) have $g = -1/2$. Each all-even cell is adjacent to $n$ one-odd cells across axis-aligned facets, so the arrangement contains at least $n\cdot\lfloor m/2\rfloor^n = \Omega(w^n)$ sign-change facets (for fixed~$n$). For $\lambda$ sufficiently large, the smooth zero set $\{f_\lambda = 0\}$ contains a distinct compact component near each such facet, and each component contributes at least $2$ to the Gauss-map count $\md\gamma_V$. Hence $\md(V) = \Omega(w^n)$.
\end{remark}


\begin{remark}\label{rem:why-not-fewnomials}
One might hope to sharpen~\eqref{eq:single-layer-bound} using a real-fewnomial bound in place of BKK, since the system~\eqref{eq:laurent-system}--\eqref{eq:laurent-system-j} has highly structured support. However, fewnomial bounds such as the Bihan--Sottile bound~\cite{bihansottile2007} are controlled by the number of monomials $|A|$, not by $\operatorname{Vol}(\conv(A))$.
\end{remark}

For networks with more than one hidden layer, we expect the single-layer bound of Proposition~\ref{prop:single-layer-degree} to extend to a fully polynomial bound in each layer width, as recorded in the conjecture below.

\begin{conjecture}\label{conj:multi-layer-degree}
  Let $F\colon \R^n \to \R^m$ be a fully connected neural network classifier with $\ell$ hidden layers of widths $n_1,\ldots,n_\ell \geq n$, logistic sigmoid activation at each hidden layer, and a linear output layer. Assume that the weight matrices $A^{(1)},\ldots,A^{(\ell)}$ are rational with common denominator $q$ and in general position, and let $L=q\max_{k,i,j}|a_{ki}^{(j)}|$ be the lattice constant, where the $a_{ki}^{(j)}$ are the entries of the $j$-th layer's weight matrix. Then each pairwise decision boundary $V_{ij} = \mathcal{Z}(F_i-F_j)$, if bounded and smooth, satisfies
  \begin{equation}\label{eq:multi-layer-bound}
    \md(V_{ij}) \leq C(n,L)\cdot \left(\prod_{i=1}^\ell n_i\right)^n,
  \end{equation}
  where $C(n,L)$ depends only on $n$ and $L$.
\end{conjecture}

The conjecture is supported by two observations. The single-layer case $\ell = 1$ is Proposition~\ref{prop:single-layer-degree}, where the exponent $n$ is sharp by Remark~\ref{rem:single-layer-lowerbound}. For piecewise-linear (ReLU) networks, which lie outside the Pfaffian framework but share the layered composition structure, the number of linear regions---and hence the number of affine pieces making up the decision boundary---is bounded by $O(\prod_j n_j^{\,n})$~\cite{montufar2014,hanin2019}, matching the form of~\eqref{eq:multi-layer-bound}. Recovering this scaling for sigmoid networks by Pfaffian methods appears to require an improvement of the Khovanski\u\i--Rolle factor $2^{s(s-1)/2}$ for chains of length $s = n_1+\cdots+n_\ell$ to one that reflects the layered structure of the chain (the ``layered Rolle'' problem; see Conjecture~\ref{conj:block-khovanskii}). A direct BKK argument on the compounded system is obstructed by the transcendentality of sigmoid composition: after a single layer of the exponential substitution used in Proposition~\ref{prop:single-layer-degree}, the deeper pre-activations involve exponentials of rational---rather than affine---functions, which are not Laurent monomials in any natural substitution.

\begin{remark}
Bianchini and Scarselli~\cite{bianchini2014complexity} obtain comparable polynomial bounds for networks with arctan activation by a different route: the arctan chain has length $2$ regardless of the number of hidden units, so the Pfaffian exponential factor is trivial. For sigmoid networks the chain length grows linearly with the number of hidden units, and a polynomial bound analogous to the arctan case is the content of Conjecture~\ref{conj:multi-layer-degree}.
\end{remark}

\subsection{Tubular neighbourhood bound for neural networks}

Combining Theorem~\ref{thm:tube_bound} with the BKK degree bound of Proposition~\ref{prop:single-layer-degree}, we obtain a tubular neighbourhood bound for single-hidden-layer sigmoid classifiers whose dependence on the width $w$ is polynomial rather than exponential.
We first extend Proposition~\ref{prop:single-layer-degree} from $\md(V)$ to the $i$-th maximal degrees $\md_i(V)$.

\begin{proposition}\label{prop:single-layer-ideg}
  Let $f\colon\R^n\to\R$ and $V = \mathcal{Z}(f)$ be as in Proposition~\ref{prop:single-layer-degree}, with lattice constant $L=q\max_{k,i}|a_{ki}|$. Then for every $0\le i\le n-1$,
  \begin{equation}\label{eq:single-layer-ideg}
    \md_i(V) \;\leq\; 2\cdot (i+1)!\,(2L+2)^{i+1}\,w^{i+1}.
  \end{equation}
\end{proposition}

\begin{proof}
  Fix an affine subspace $L'\in\cE^n_{i+1}$ such that $V\cap L'$ is transverse. By Remark~\ref{re:affine}, choose an orthogonal transformation $h\in O(n)$ such that $h(L')=\{x\in\R^n\colon x_{i+2}=\cdots=x_n=0\}$, and write $\tilde f(u)=f(h^{-1}(u,0))$ for $u\in\R^{i+1}$. Then
  \begin{equation*}
    \tilde f(u) = c_0 + \sum_{k=1}^w d_k\,\sigma(\tilde a_k^{\trans}u + b_k),
  \end{equation*}
  where $\tilde a_k\in\R^{1+i}$ is the projection of $a_k$ onto the first $i+1$ coordinates in the rotated frame. The entries of $\tilde a_k$ need not be rational.

  For each $k$, let $\tilde a_k'\in(1/q)\Z^{1+i}$ be the coordinatewise round of $\tilde a_k$ to the nearest multiple of $1/q$. Then $\tilde a_k'$ has common denominator $q$ and $\|\tilde a_k'-\tilde a_k\|_\infty\le 1/(2q)$; since $\|\tilde a_k\|_\infty\le\|a_k\|_\infty\le L/q$ by orthogonal projection, the rational lattice constant for the approximants is $L^*:=q\max_k\|\tilde a_k'\|_\infty\le L+1$. Let
  \begin{equation*}
    \tilde f'(u) \;=\; c_0+\sum_{k=1}^w d_k\,\sigma(\tilde a_k'^{\trans}u+b_k).
  \end{equation*}
  Applying Proposition~\ref{prop:single-layer-degree} in dimension $i+1$ to $\tilde f'$ with lattice constant $L^*\le L+1$ gives
  \begin{equation}\label{eq:ideg-approx}
    \md(\mathcal{Z}(\tilde f')) \;\leq\; 2\cdot(i+1)!\,(2L+2)^{i+1}\,w^{i+1},
  \end{equation}
  after an arbitrarily small further perturbation of $(\tilde a_k',b_k,c_0,d_k)$ to restore general position and the Sard-type non-degeneracy hypotheses of Proposition~\ref{prop:single-layer-degree}, which hold outside a measure-zero subset of parameter space.

  By repeating the construction with $q$ replaced by $Nq$ for $N=1,2,\ldots$, the approximants $\tilde f'^{(N)}$ converge to $\tilde f$ in $C^2_{\mathrm{loc}}$ as $N\to\infty$. For any regular value $v$ of the Gauss map of $V\cap L'$, the corresponding real solutions of the Gauss-map system~\eqref{eq:single-layer-system} for $\tilde f$ are non-degenerate, and by the implicit function theorem each persists as a unique non-degenerate real solution of the perturbed system for all $N$ sufficiently large. Upper semicontinuity of the intersection number under compact perturbations~\cite[Ch.~3]{hirsch1976differential,guillemin2010differential} gives
  \begin{equation*}
    \md(V,L') \;\leq\; \liminf_{N\to\infty}\md(\mathcal{Z}(\tilde f'^{(N)})) \;\leq\; 2\cdot(i+1)!\,(2L+2)^{i+1}\,w^{i+1}.
  \end{equation*}
  Taking the supremum over $L'\in\cE^n_{i+1}$ yields~\eqref{eq:single-layer-ideg}.
\end{proof}

In particular, using $w\ge 1$ and $i+1\le n$, Proposition~\ref{prop:single-layer-ideg} gives the uniform bound $\md_i(V)\le 2\cdot n!\,(2L+2)^n\,w^n$ for $0\le i\le n-1$. We now state the main result of this section: the analogue of Theorem~\ref{thm:prob_bound} for the decision boundary of a single-hidden-layer sigmoid network. We restrict to the case $V\subset B(p,\sigma)$, which eliminates the sphere-boundary term $\partial M$ from the proof of Theorem~\ref{thm:prob_bound}; this assumption is natural in the robustness setting of Section~\ref{sec:robustness}, where the data distribution is supported in a bounded region and $\sigma$ exceeds its diameter. The obstruction to dropping this hypothesis is recorded in Remark~\ref{rem:sphere-obstruction} below.

\begin{theorem}\label{thm:prob_bound_single_layer}
  Let $f\colon\R^n\to\R$ be as in Proposition~\ref{prop:single-layer-degree} and set $V=\mathcal{Z}(f)$. Assume that $\nabla f$ is non-vanishing on $V$ and that $V\subset B(p,\sigma)$. Let $X$ be uniformly distributed on $B(p,\sigma)$. Then
  \begin{equation}\label{eq:single-layer-prob-bound}
    \prob\{d(X,V)\le\varepsilon\}
    \;\leq\; C_n(L)\cdot w^n\left[\left(1+\frac{\varepsilon}{\sigma}\right)^n - 1\right],
  \end{equation}
  where $C_n(L) \le 4\cdot n!\,(2L+2)^n$ depends only on $n$ and $L$.
\end{theorem}

\begin{proof}
  Since $V\subset B(p,\sigma)$ is compact, Theorem~\ref{thm:tube_bound} applies to $V$ as a codimension-$1$ submanifold inside the ball $B(p,\sigma)$, giving
  \begin{equation*}
    \vol T(V,\varepsilon)
    \;\leq\; 2\omega_n\sigma^n\sum_{i=0}^{n-1}\binom{n}{i+1}\md_i(V)\left(\frac{\varepsilon}{\sigma}\right)^{i+1}.
  \end{equation*}
  Substituting the uniform bound $\md_i(V)\le 2\cdot n!\,(2L+2)^n\,w^n$ from Proposition~\ref{prop:single-layer-ideg} and using the binomial identity $\sum_{j=1}^n\binom{n}{j}t^j=(1+t)^n-1$ with $t=\varepsilon/\sigma$,
  \begin{equation*}
    \vol T(V,\varepsilon)
    \;\leq\; 4\omega_n\sigma^n\cdot n!\,(2L+2)^n\,w^n\left[\left(1+\frac{\varepsilon}{\sigma}\right)^n-1\right].
  \end{equation*}
  The inclusion $T(V,\varepsilon)\cap B(p,\sigma)\subseteq T(V,\varepsilon)$ and division by $\vol B(p,\sigma)=\omega_n\sigma^n$ give~\eqref{eq:single-layer-prob-bound} with $C_n(L)=4\cdot n!\,(2L+2)^n$.
\end{proof}

\begin{corollary}\label{cor:single-layer-tight-prob-bound}
  Under the hypotheses of Theorem~\ref{thm:prob_bound_single_layer},
  \begin{equation*}
    \prob\{d(X,V)\le\varepsilon\}
    \;\leq\; n\,C_n(L)\cdot w^n\cdot\frac{\varepsilon}{\sigma}\left(1+\frac{\varepsilon}{\sigma}\right)^{n-1}.
  \end{equation*}
  In particular, for $\varepsilon\le\sigma$,
  \begin{equation*}
    \prob\{d(X,V)\le\varepsilon\} \;\leq\; 2^{n-1}\,n\,C_n(L)\cdot w^n\cdot\frac{\varepsilon}{\sigma}.
  \end{equation*}
\end{corollary}

\begin{proof}
  By the mean-value theorem, for $t\ge 0$ there exists $\xi\in[1,1+t]$ with $(1+t)^n - 1 = n\,\xi^{n-1}\,t \le n\,t\,(1+t)^{n-1}$. Applied to~\eqref{eq:single-layer-prob-bound} with $t=\varepsilon/\sigma$ this gives the first assertion; for $\varepsilon\le\sigma$, $(1+\varepsilon/\sigma)^{n-1}\le 2^{n-1}$.
\end{proof}

The linear-in-$\varepsilon/\sigma$ scaling of Corollary~\ref{cor:single-layer-tight-prob-bound} is essential for deriving condition-number tail bounds of order $O(1/t)$ in Section~\ref{sec:robustness}; the corresponding statement in the general Pfaffian setting is Corollary~\ref{cor:tight_prob_bound}, whose exponential prefactor $2^{h(h-1)/2}$ is here replaced by the polynomial $w^n$ for single-hidden-layer networks.

\begin{remark}[Sphere obstruction]\label{rem:sphere-obstruction}
  The hypothesis $V\subset B(p,\sigma)$ in Theorem~\ref{thm:prob_bound_single_layer} eliminates the sphere-boundary contribution $\partial M = V\cap S^{n-1}(p,\sigma+\varepsilon)$ that appears in the proof of Theorem~\ref{thm:prob_bound}. Without it, one needs to bound the maximal degrees of $\partial M$, a codimension-$2$ intersection of a Pfaffian hypersurface with a polynomial sphere. The Khovanski\u\i{} complete-intersection bound~\eqref{eq:ideg} applied to $\partial M$ with $(\alpha,\beta,s)=(2,1,w)$ and $c=2$ yields a correction term of order $2^{w(w-1)/2}\cdot\mathrm{poly}(n,w)$, exponential in $w$ and dominating the BKK main term~\eqref{eq:single-layer-prob-bound}. A polynomial-in-$w$ replacement would require either a Pfaffian--algebraic B\'ezout-type inequality of the form $\md(V\cap S)\le\md(V)\cdot\deg(S)$ for Pfaffian hypersurfaces $V$ and polynomial hypersurfaces $S$, or a BKK-style global argument that accommodates mixed Laurent-polynomial and transcendental components: under the exponential substitution of Proposition~\ref{prop:single-layer-degree}, the sphere equation $\|x-p\|^2=r^2$ becomes $\sum_j(q\log y_j+p_j)^2=r^2$, which is transcendental in $y$. This is a weaker form of the layered-Rolle obstruction discussed in Section~\ref{subsec:improved-nn} (Conjecture~\ref{conj:block-khovanskii}).
\end{remark}
